---
title: "Assignment 3: Global CO~2~ Emissions"
author: "Xiyu Zhang"
date: "2022-09-23"
output: html_document
---

<!-- Please knit this right away! The additional information for (html and css)
are provided to help you build on your current R Markdown toolbox. You will not
be responsible for learning these additional tools at this point. -->

<!-- The head tag, <head>, is a container for metadata in an html document. 
We can use it to define styles and do lots of other cool things things. -->
<head>
<!-- The link tag, <link>, creates a relationship between this file and
an external source. Here, I'm linking font-awesome so I can include a 
couple of icons that I like (user-secret and user-circle) -->
<link 
  rel="stylesheet" 
  href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">
<!-- Here I'm linking to a javascript file that will give us access to font awesome -->
  <script src="https://kit.fontawesome.com/03064bdd6c.js" crossorigin="anonymous"></script>
</head>
<!-- Note in the above that some html tags, like <head> need to be closed 
with </head>. Others, like <link> don't need a closing tag. -->

```{css, echo = FALSE}
/* We can use the body tag to change the default text size of the whole document. */
  
body {
  font-size: 18px
}

/* We can increase the space between our list items */
  
ul, li {
  padding-top: 0.25em;
}

pre {
  margin-top: 2em;
  margin-bottom: 2em;
}

/* I can add custom classes with a period symbol. */

.score {
  font-weight: bold;
  color: red;
  background-color: #ffff80;
}

/* I create an custom class for user secrets. Have a look at the area surrounding my spy icons below to see what this does. Also, do you notice any similarity to your ggPlot theme elements? */
  
.mysecret {
  border: 1px solid #999999;
  border-style: solid; 
  padding: 20px; 
  border-radius: 25px;
  margin-top: 15px;
  margin-bottom: 15px;
}

/* Here, I add a custom class for area surrounding the graded portions of this assignment.*/

.now_you {
  border-left: 8px solid #292093;
  background-color: #e6ffff;
  padding: 16px;
  margin-top: 15px;
  margin-bottom: 15px;
}

/* Here I will add some custom css on my font awesome icons*/
  
.fas {
  font-size:36px
}
  
.subscore {
  font-size: 14px;
  color: red;
}

/* I can modify a current class like the below. To see what this does,
have a look at where I am using block quotes in the document.*/
  
blockquote {
  font-size: 18px;
  border-left: 8px solid #292093;
  background-color: #e6ffff;
  padding-left: 16px;
  padding-right: 16px;
  margin-top: 15px;
  margin-bottom: 15px;
}

/* Accordian style */
.accordion {
  background-color: #e6f0ff;
  color: #000000;
  cursor: pointer;
  padding: 18px;
  width: 100%;
  border-style: solid;
  border-width: 2px;
  border-color: #ffffff;
  text-align: left;
  outline: none;
  font-size: 15px;
  transition: 0.4s;
}

.active, .accordion:hover {
  background-color: #80b3ff; 
  color: #ffffff;
  font-weight: bold;
}

.panel {
  padding: 0 18px;
  display: none;
  background-color: #eeffff;
  overflow: hidden;
}
```

<!-- Here is the setup for knitr. I had to include tidyverse so I could
include a tidyverse example with my Important warning statement below-->

```{r knitr_setup, include = FALSE}
options(knitr.table.format = "html")

knitr::opts_chunk$set(error = TRUE)

library(tidyverse)
```

<hr>

## Overview

In this exercise, you will explore data compiled by the <a href = "https://databank.worldbank.org/home.aspx" target = "_blank">World Bank DataBank</a> that provides information on the emissions of greenhouse gases and the populations of countries around the world. 

**Please note: For R Markdown to work properly this file must be placed in the top level of your R Studio project folder!**

Please start by knitting this file. You can use the html file that is generated as a handy reference (it's much easier to read than the markdown file).

You will be working in two files as you complete this assignment, your own new script file, and this R markdown file I'm providing, which you'll be editing and adding to. You will submit this R Markdown file (modified as described below) and your R script file.

<!-- A div is a container for html. I assign it to the class "my secret that
we created above -->
<div class = "mysecret">
<!-- The i tag is commonly used to create italic text, but it can also be 
used to create icons -->
<i class="fas fa-user-secret"></i> When I create an R Markdown file to communicate a coding process, I usually work in a .R script file and copy-and-paste the code into the R Markdown document at the end.
</div>

<hr>

## Grading

The points allotted for each question are provided in highlighted red bold text (e.g., <span class = "score">[1.0]</span>) within the question itself. When applicable, total points for a question may represent the sum of individually graded components, which are provided in red text (e.g., <span class = "subscore">[1.0]</span>).

**Points may be deducted from each question's total:** 

* Use only functions other than those listed in "Functions that you may use in this assignment" [50%]
* Do not include any assignments other than those specified in the question [50%]
* Unless specified otherwise, each question must be answered in a single piped statement [50%]
* Include a comment at the start of each code block [25%]
* Include a comment for each line of code that modifies a dataset [10% per violation]
* Ensure that all code is properly indented [10% per violation]
* Ensure that your code follows modern coding conventions (see *Best practices in modern R coding*) [10% per violation]

*Note: The maximum deduction is the total points value for a given question*

You must ensure that your R Markdown document **runs out-of-the-box** -- in other words, the document will knit without error. Some tips for doing so:

* Do not maintain your project folder in a location that is backed up by an online directory (e.g., Dropbox, iCloud)
* Ensure that all file paths are equivalent to those in this document (e.g., in `source()` or `read_csv()`)
* Do not use `setwd()` in your code (***Never** use `setwd()`!*)
* If one of your code chunks contains an error, include `eval = FALSE` in the options for that chunk
* Prior to submission, clear all objects from your global environment prior to knitting your R Markdown document

<span style = "color: red; font-size: 36px"><i class="fas fa-exclamation-triangle"></i> **Important**</span> 

To receive full credit for a given question, please ensure that the result you provide **only** answers the given question. For example, in the table below, if you were asked to determine the value in column `b` where column `a` is equal to "hello" you would provide the following code:

```{r example}
tibble(
  a = c('hello', 'world'),
  b = 1:2) %>% 
  filter(a == 'hello') %>% 
  pull(b)
```

<hr>


<div>
<button class="accordion">Functions that you may use in this assignment</button>
<div class="panel">

In this assignment, you may use **only** the following R functions (*Note: If you are unclear on what a given function does, use `?` to view the help file!*):

* base::<-
* base::=
* base::!
* base::(
* base::)
* base::%in%
* base::c
* base::library
* dplyr::full_join
* dplyr::inner_join
* dplyr::mutate
* dplyr::pivot_longer
* dplyr::rename
* dplyr::select
* dplyr::transmute
* ggplot2::aes
* ggplot2::facet_wrap
* ggplot2::geom_point
* ggplot2::geom_line
* ggplot2::ggplot
* ggplot2::theme
* magrittr::%>%
* readr::read_csv

*Note: The packages dplyr, ggplot2, magrittr, readr, rlang, and tibble are all part of the tidyverse and are loaded with `library(tidyverse)`.*

</div>
</div>

<hr>

## Getting started

As always, please ensure that you are starting with a clean session.

Open R Studio. Remember that it's best practice to start with a clean R Studio session. Please do the following before continuing:

1. If there are any script files open in your source pane. Close them. If any of the file titles are blue, save them prior to closing.

1. In the *Environment* tab of your **workspace pane**, ensure that your **global environment** is empty. If it is not, click the *broom* to remove all objects.

1. In the *History* tab of your **workspace pane**, ensure that your history is empty. If it is not, click the *broom* to remove your history.

<div class = "now_you">
<i class="fas fa-user-circle"></i> 
<hr>
1\. <span class = "score">[0.5]</span> Save this document with the name: `data_sci/output/assignment_3_Evans_Brian.Rmd` (replacing my name with yours).
</div>

### Create a script file:

1. Open a new script file.

1. Save the file right away.

1. Add a new code section and call it "setup"
<br>

<div class = "now_you">
<i class="fas fa-user-circle"></i> 
<hr>
2\. <span class = "score">[0.25]</span> After a space between your section break and the code, include and run the following:

```{r q2, eval = TRUE, results = 'hide', message = FALSE}
# Load libraries:

library(tidyverse)
```
</div>
<hr>

## Read in the data

I've provided three data sets:

* Country-level information: `world_bank_countries.csv` *Note: I have cleaned this dataset for you* 
* Population data, by country, from 1960 through 2020: `API_SP.POP.TOTL_DS2_en_csv_v2_2763937.csv`
* CO~2~ emission data, by country, from 1960 through 2020: `API_EN.ATM.CO2E.PC_DS2_en_csv_v2_2764620.csv`

*Note: The data set `world_bank_countries.csv` has been cleaned for you. The population and CO~2~ data sets are in their raw form (including the names).*

In your script file, create a new code section for reading in the data.

<div class = "now_you">
<i class="fas fa-user-circle"></i> 
<hr>
3\. <span class = "score">[0.5]</span>  Read in the Country-level information and store it in your global environment by assigning the name `countries`. The files should be located in the folder `data/processed`. Be sure to include a comment that describes which file you are reading in.

```{r q3}

# Paste your code here!

```
</div>

Reading and processing the raw data will be somewhat more challenging.

Run the script below to read in the population data and assign the name `world_pop` to the object (*Note: You must also execute this code block in your script file!*):

```{r read_world_pop, message = FALSE, warning = FALSE}
# Read in population data:

world_pop <-
  read_csv(
    'data/raw/API_SP.POP.TOTL_DS2_en_csv_v2_2763937.csv', 
    skip = 3)
```

I usually do initial data cleaning as I read in a file. I'm going to demonstrate my reading and cleaning operation using the population data.

Let's look at the data first (do so in your Console pane):

```{r world_pop_print}
world_pop
```

We can see that the data are super wide. These data are certainly breaking the rule that each observation forms a row. I can fix that with `pivot_longer()`:

```{r world_pop_pivot, message = FALSE, warning = FALSE}
# Read in population data:

world_pop <-
  read_csv(
    'data/raw/API_SP.POP.TOTL_DS2_en_csv_v2_2763937.csv', 
    skip = 3) %>% 
  
  # Reshape year columnns into rows:
  
  pivot_longer(
    names_to = 'year',
    values_to = 'population',
    `1960`:`2020`)
```

Have a look at the data (do so in your Console pane):

```{r world_pop_pivot_print}
world_pop
```

Column names that have spaces are a bear to work with. I'm going to change the name of the ``Country Code`` column using `rename()`:

```{r world_pop_rename, message = FALSE, warning = FALSE}
# Read in population data:

world_pop <-
  read_csv(
    'data/raw/API_SP.POP.TOTL_DS2_en_csv_v2_2763937.csv', 
    skip = 3) %>% 
  
  # Reshape year columnns into rows:
  
  pivot_longer(
    names_to = 'year',
    values_to = 'population',
    `1960`:`2020`) %>% 
  
  # Rename columns with a space:
  
  rename(country_code = `Country Code`)
```

*Note: Why did I include "skip = 3" as one of my arguments? You might want to look at what happens when you don't include a skip argument. If this is confusing, please revisit our lecture or exercise on reading and writing data.*

Have a look at the data again (do so in your Console pane):

```{r world_pop_rename_print}
world_pop
```

And finally, I'm going to grab my columns of interest, which here are `country_code`, `year`, and `population`:

```{r world_pop_select, message = FALSE, warning = FALSE}
# Read in population data:

world_pop <-
  read_csv(
    'data/raw/API_SP.POP.TOTL_DS2_en_csv_v2_2763937.csv', 
    skip = 3) %>% 
  
  # Reshape year columnns into rows:
  
  pivot_longer(
    names_to = 'year',
    values_to = 'population',
    `1960`:`2020`) %>% 
  
  # Rename columns with a space:
  
  rename(country_code = `Country Code`) %>% 
  
  # Select columns of interest:
  
  select(country_code, year, population)
```

Have a look at the data again:

```{r world_pop_select_print}
world_pop
```

In your script file, replace the code section that read in and assigned `world_pop` with the code section above that: 1) read in the file, 2) pivoted the data, 3) renamed the country column, and 4) selected the columns of interest.

<div class = "now_you">
<i class="fas fa-user-circle"></i> 
<hr>
4\. <span class = "score">[1.0]</span> Read in the CO~2~ data <span class = "subscore">[0.25]</span> and store it in your global environment by assigning the name `co2` <span class = "subscore">[0.25]</span>. In doing so, modify the code section that read in the CO~2~ data such that the resultant tibble contains (only) the column names `country_code`, `year`, and `co2` (as above) <span class = "subscore">[0.5]</span>.

```{r q4}

# Paste your code here!

```
</div>
<hr>

## Normalize and clean the data

In your script file, create a new code section and call it "normalize and clean data".

We can see that the data are not tidy. Recall the tidy rule that every level of observation forms a table. The population table and the CO~2~ table are the same level of observation -- each characterize some attribute of the country on a given year. As such, the data should be stored in the same table.  We'll need to join the tables together.

Type `?join` in your console. Which join should you use? *Note: you'll have to be careful here, because the two tibbles contain two common columns that define the observation! This is what is known as a **compound key**.*

<div class = "now_you">
<i class="fas fa-user-circle"></i> 
<hr>
5\. <span class = "score">[1.25]</span> Add a comment that *briefly* states which aspect of the data you are normalizing <span class = "subscore">[0.5]</span> and join the population and CO~2~ tibbles <span class = "subscore">[0.5]</span>. Assign the name, `population_co2` to the resultant object<span class = "subscore">[0.25]</span>. 

```{r q5}

# Paste your code here!

```

If you type `str(world_pop)` in our console pane, we can see that year is stored as a character vector. It should be numeric. 

6\. <span class = "score">[0.5]</span> Modify your above code section to change the year column to numeric. 

```{r q6}

# Paste your code here!

```
</div>
<hr>

## Subset and plot the data

<span class = "score">[0.5]</span> In your script file, create a new code section and call it "plot population density and co2 emissions in Africa".

Our next goal is to explore the change in population density in African countries, over time. For this, we will need the size of each country of interest -- that is stored in the `countries` data object. 

<div class = "now_you">
<i class="fas fa-user-circle"></i> 
<hr>
7\. <span class = "score">[1.5]</span> Subset `population_co2` such that:

* <span class = "subscore">[1.0]</span>It only includes countries that are in Africa;
* <span class = "subscore">[0.5]</span>The resultant object includes only the variables `country_name`, `land_area`, `population`, and `co2`. 

```{r q7}

# Paste your code here!

```

Now we want to calculate population density, which is the population per unit area (`population/land_area`), and subset the data to our columns of interest. 

8\. <span class = "score">[1.5]</span>  Modify the code section above:

* <span class = "subscore">[0.5]</span> Calculate the population density of each country and assign the vector to your data frame with the name `population_density`;
* <span class = "subscore">[0.75]</span> Subset the data to the columns  `country_name`, `population_density`, `co2` and `year` (as parsimoniously as possible);
* <span class = "subscore">[0.25]</span> Assign the object to your global environment as `africa_population_co2`.

```{r q8}

# Paste your code here!

```


9\. <span class = "score">[1.0]</span> Use ggPlot to visualize the change in population in Africa over time: 

* <span class = "subscore">[0.5]</span> Map `year` to your x-axis, `population_density` to your y-axis, and `country_name` to color;
* <span class = "subscore">[0.5]</span> Add point **and** line geometries to the plot and include at least one modification to the default theme.

```{r q9}
# Paste your code here!

```

10\. <span class = "score">[2.0]</span> Let's now make a faceted plot. One facet will be population density and the other facet will be CO~2~ emissions. To do so, we'll need to put both of our measurements in the same column. 

* <span class = "subscore">[1.0]</span> Add a line of code that reshapes the data such that: 
    + The column names `population_density` and `co2` are stored in a single column called `variable`
    + The values for `population_density` and `co2` are stored in a single column called `value`.
* <span class = "subscore">[1.0]</span> Use `facet_wrap` to create a 2-facet plot, where one facet represents `population_density` and the other facet `co2`. 

<hr>

```{r q10}

# Paste your code here!

```
</div>

## Submission

Please save your R markdown (.Rmd) and script (.R) files.

**Only submit the rmd** file!

<hr>

<script>
var acc = document.getElementsByClassName("accordion");
var i;

for (i = 0; i < acc.length; i++) {
  acc[i].addEventListener("click", function() {
    this.classList.toggle("active");
    var panel = this.nextElementSibling;
    if (panel.style.display === "block") {
      panel.style.display = "none";
    } else {
      panel.style.display = "block";
    }
  });
}
</script>

