---
title: "Assignment 6: A mountain holiday for Alan"
author: "Xiyu Zhang"
date: "2022-11-19"
output: html_document
---

<!-- Please knit this right away! The additional information for (html and css)
are provided to help you build on your current R Markdown toolbox. You will not
be responsible for learning these additional tools at this point. -->

<!-- The head tag, <head>, is a container for metadata in an html document. 
We can use it to define styles and do lots of other cool things things. -->
<head>
<!-- The link tag, <link>, creates a relationship between this file and
an external source. Here, I'm linking font-awesome so I can include a 
couple of icons that I like (user-secret and user-circle) -->
<link 
rel="stylesheet" 
href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">
<!-- Here I'm linking to a javascript file that will give us access to font awesome -->
<script src="https://kit.fontawesome.com/03064bdd6c.js" crossorigin="anonymous"></script>
</head>
<!-- Note in the above that some html tags, like <head> need to be closed 
with </head>. Others, like <link> don't need a closing tag. -->

```{css, echo = FALSE}
/* We can use the body tag to change the default text size of the whole document. */

body {
font-size: 18px
}

/* We can increase the space between our list items */

ul, li {
padding-top: 0.25em;
}

pre {
margin-top: 2em;
margin-bottom: 2em;
}

/* I can add custom classes with a period symbol. */

.score {
font-weight: bold;
color: red;
background-color: #ffff80;
}

/* I create an custom class for user secrets. Have a look at the area surrounding my spy icons below to see what this does. Also, do you notice any similarity to your ggPlot theme elements? */

.mysecret {
border: 1px solid #999999;
border-style: solid; 
padding: 20px; 
border-radius: 25px;
margin-top: 15px;
margin-bottom: 15px;
}

/* Here, I add a custom class for area surrounding the graded portions of this assignment.*/

.now_you {
border-left: 8px solid #292093;
background-color: #e6ffff;
padding: 16px;
margin-top: 15px;
margin-bottom: 15px;
}

/* Here I will add some custom css on my font awesome icons*/

.fas {
font-size:36px
}

.subscore {
font-size: 14px;
color: red;
}

/* I can modify a current class like the below. To see what this does,
have a look at where I am using block quotes in the document.*/

blockquote {
font-size: 18px;
border-left: 8px solid #292093;
background-color: #e6ffff;
padding-left: 16px;
padding-right: 16px;
margin-top: 15px;
margin-bottom: 15px;
}

/* Accordion style */
.accordion {
background-color: #e6f0ff;
color: #000000;
cursor: pointer;
padding: 18px;
width: 100%;
border-style: solid;
border-width: 2px;
border-color: #ffffff;
text-align: left;
outline: none;
font-size: 15px;
transition: 0.4s;
}

.active, .accordion:hover {
background-color: #80b3ff; 
color: #ffffff;
font-weight: bold;
}

.panel {
padding: 0 18px;
display: none;
background-color: #eeffff;
overflow: hidden;
}
```

<!-- Here is the setup for knitr. I had to include tidyverse so I could
include a tidyverse example with my Important warning statement below-->

```{r knitr_setup, include = FALSE}
options(knitr.table.format = "html")

knitr::opts_chunk$set(error = TRUE)

library(tidyverse)
```

<hr>

## Overview

<div class = "row">

<figure 
style = 
"float: right; 
padding-left: 1em; 
padding-bottom: 1em; 
padding-right: 1em; 
width: 60%; 
height: auto;">
<img  
src = "https://www.nps.gov/articles/000/images/black-bear-Glacier-Bay-NPS-resized.jpg?maxwidth=1200&autorotate=false&quality=78&format=webp"
alt="Bear and dandelions">
</img>
<figcaption style = "font-size: 0.75em;">*A frighteningly "goofy" American Black Bear munching dandelions*</figcaption>
</figure>

<div style = "padding-left: 1em;">

Being a student takes a lot of energy and patience --
especially when your teacher is someone as annoying as me. At the end of every
semester Alan McQuinn likes to take a short mountain holiday to regain his 
strength. This semester, after dealing with me for more than four months, Alan
really needs it. Right after the semester ends, Alan is going to get the rest
and relaxation he deserves in Pisgah National Forest -- a stunningly beautiful
spot in the Blue Ridge Mountains near Asheville, North Carolina. Let's send him
on an awesome, rejuvenating, Brian-free, vacation! 

In this exercise, you'll help Alan find the perfect spots to stargaze, stay 
warm, search for waterfalls, get some quiet time, and, most importantly, 
avoid goofy bears!

</div>

</div>

<!-- A div is a container for html. I assign it to the class "my secret that
we created above -->
<div class = "mysecret">
<!-- The i tag is commonly used to create italic text, but it can also be 
used to create icons -->
<i class="fas fa-user-secret"></i> When I create an R Markdown file to 
communicate a coding process, I usually work in a .R script file and 
copy-and-paste the code into the R Markdown document at the end.
</div>

<hr>

<span style = "color: red; font-size: 36px">
<i class="fas fa-exclamation-triangle"></i> **Important**</span> 

To receive full credit for a given question, please ensure that the result you
provide **only** answers the given question. For example, in the table below, 
if you were asked to determine the value in column `b` where column `a` is equal
to "hello" you would provide the following code:

```{r example}
tibble(
  a = c('hello', 'world'),
  b = 1:2) %>% 
  filter(a == 'hello') %>% 
  pull(b)
```

<hr>

## Grading

The points allotted for each question are provided in highlighted red bold text
(e.g., <span class = "score">[1.0]</span>) within the question itself. When
applicable, total points for a question may represent the sum of individually
graded components, which are provided in red text 
(e.g., <span class = "subscore">[1.0]</span>).

**Points may be deducted from each question's total:** 

* Use only functions other than those listed in "Functions that you may use in 
this assignment" [50%]
* Do not include any assignments other than those specified in the question
[50%]
* Unless specified otherwise, each question must be answered in a single piped 
statement [50%]
* Include a comment at the start of each code block [25%]
* Include a comment for each line of code that modifies a dataset [10% per 
violation]
* Ensure that all code is properly indented [10% per violation]
* Ensure that your code follows modern coding conventions (see *Best practices
in modern R coding*) [10% per violation]

*Note: The maximum deduction is the total points value for a given question*

You must ensure that your R Markdown document **runs out-of-the-box** -- in 
other words, the document will knit without error. Some tips for doing so:

* Do not maintain your project folder in a location that is backed up by an 
online directory (e.g., Dropbox, iCloud)
* Ensure that all file paths are equivalent to those in this document (e.g., in
`source()` or `read_csv()`)
* Do not use `setwd()` in your code (***Never** use `setwd()`!*)
* If one of your code chunks contains an error, include `eval = FALSE` in the
options for that chunk
* Prior to submission, clear all objects from your global environment prior to
knitting your R Markdown document
<hr>


<div>
<button class="accordion">Functions that you may use in this assignment</button>
<div class="panel">

In this assignment, you may use **only** the following R functions (*Note: If
you are unclear on what a given function does, use `?` to view the help file!*):

* base:: ::
* base::<-
* base::=
* base::<
* base::>
* base::<=
* base::>=
* base::!
* base::+
* base::-
* base::*
* base::/
* base::$
* base::!
* base::~
* base::(
* base::)
* base::{
* base::}
* base::as.matrix
* base::c
* base::is.na
* base::library
* base::names
* base::rm
* base::source
* dplyr::if_else
* dplyr::filter
* dplyr::mutate
* dplyr::pull
* dplyr::slice
* dplyr::transmute
* grDevices::gray.colors
* grDevices::terrain.colors
* magrittr::%>%
* purrr::keep
* purrr::map
* sf::st_as_sf
* sf::st_coordinates
* sf::st_crs
* sf::st_set_crs
* spatstat::as.owin
* spatstat::ppp
* stats::density
* stars::st_as_stars
* stringr::str_detect
* terra::as.points
* terra::classify
* terra::crop
* terra::distance
* terra::ext
* terra::extract
* terra::mask
* terra::rast
* terra::resample
* terra::shade
* terra::terrain
* terra::vect
* tibble::tibble
* tibble::tribble
* tmap ... any tmap function you would like!



*Note: The packages dplyr, ggplot2, magrittr, readr, rlang, and tibble are all
part of the tidyverse and are loaded with `library(tidyverse)`.*

</div>
</div>

<hr>

## Getting started

Open R Studio. Please ensure that the Global Options for R Markdown are set to
"Show output preview in Viewer Pane". This will allow you to view the knitted 
document (i.e., the html version of this file) right next to your code.

As always, please ensure that you are starting with a clean session. Please do
the following before continuing:

* If there are any script files open in your source pane. Close them. If any of
the file titles are blue, save them prior to closing.

* In the *Environment* tab of your **workspace pane**, ensure that your **global
environment** is empty. If it is not, click the *broom* to remove all objects.

* In the *History* tab of your **workspace pane**, ensure that your history is
empty. If it is not, click the *broom* to remove your history.

<div class = "now_you">
<i class="fas fa-user-circle"></i> 

<hr>

1\. <span class = "score">[0.5]</span> Save and knit this document:

* Replace my name in the YAML header with yours <span class = "subscore">
[0.1]</span>
* Add the current date in the YAML header <span class = "subscore">[0.1]</span>
* Save the *.rmd* file in the **output** folder of your project as (*but replace
my name with yours*): `assignment_6_Evans_Brian.rmd` <span class = "subscore">
[0.3]</span>
</div>
<hr>

## The setup

**All** of my scripts start with a setup section where I load libraries, set
session options, conduct initial data loading, and early processing steps.

<div class = "now_you">
<i class="fas fa-user-circle"></i> 
<hr>
2\. <span class = "score">[0.5]</span>  Setup your session:

* <span class = "subscore">[0.15]</span> Load the `spatstat`, `sf`, `tmap`, and
`tidyverse` libraries.

```{r question_2a, results = 'hide', message = FALSE, warning = FALSE}

# Load libraries

library(spatstat)
library(sf)
library(tmap)
library(tidyverse)

```


* <span class = "subscore">[0.15]</span> Set the tmap mode to "view" for the
entire document:

```{r question_2b, results = 'hide', message = FALSE, warning = FALSE}

# Set the tmap mode to 'view'

tmap_mode('view')

```

* <span class = "subscore">[0.2]</span> Load the data using the source script
`source_script_holiday_Alan.R`:

```{r question_2c, results = 'hide', message = FALSE, warning = FALSE}

# Load the data using the source script

source('scripts/source_script_holiday_Alan.R')

```


3\. <span class = "score">[0.5]</span> Using iteration, subset each raster file
in `rasters_start` to the extent below <span class = "subscore">[0.3]</span> and
assign the results to your global environment as a list object with the name
"rasters_reduced" <span class = "subscore">[0.1]</span>:

| Coordinate        |   Value |
|:-----------------:| -------:|
| Minimum longitude | 1172400 |
| Maximum longitude | 1211000 |
| Minimum latitude  | 1433000 |
| Maximum latitude  | 1472000 |


```{r question_3a, results = 'hide', message = FALSE}

# Subset raster files to specific extent by a loop

rasters_reduced <-
  rasters_start %>%
  map(
    ~ terra::crop(.x,
                  terra::ext(1172400, 1211000, 1433000, 1472000)))

```

* <span class = "subscore">[0.1]</span> Remove the original raster files from 
the global environment:

```{r question_3b, results = 'hide', message = FALSE}

# Remove the original raster files

rm(rasters_start)

```

4\. <span class = "score">[1.0]</span> Processing climate variables:

* <span class = "subscore">[0.25]</span> Using iteration, set the resolution of
the `rasters_reduced$precip` and `rasters_reduced$tmin` rasters to the 
resolution of rasters_reduced$canopy;
* <span class = "subscore">[0.25]</span> Convert the modified climate rasters to
a raster stack;
* <span class = "subscore">[0.25]</span> Assign the raster stack to the global
environment with the name "climate";

```{r 4_a-c, results = FALSE}

# Processing climate variables:

climate <- 
  list(rasters_reduced$precip,
       rasters_reduced$tmin) %>% 
  
  # using iteration to reset the resolution of the two rasters
  
  map(
    ~ terra::resample(.x,
                      
                      # reset the resolution as same as the canopy raster
                      
                      rasters_reduced$canopy)) %>% 
  
  # convert the modified climate rasters to a raster stack
  
  terra::rast()

```

* <span class = "subscore">[0.25]</span> Briefly describe why the climate raster
stack should not be used analytically:

```{r 4d}

# Because initially the resolution of these rasters are much larger. By changing
# their resolution to the same as rasters_reduced$canopy, we changed the 
# resolution from a larger grain to a smaller grain, which means we are creating
# some data don't exist initially. Thus, this raster stack should not be used 
# analytically.

```

</div>
<hr>

## Helping out ol' Alan

With our processing complete, we're now ready to get to the fun part --
extracting meaningful data!

*Aside: I know that at least a few of you (Xavier? Rui? Anyone?) are saying to
yourself, "but Brian, the processing **is** the fun part". If so, I get it, but
we've at least got to pretend otherwise ...*

<div class = "now_you">
<i class="fas fa-user-circle"></i> 
<hr>
5\. <span class = "score">[1.0]</span> Alan sure is excited about his trip! Make
him a nice map of the topography of Western North Carolina. Use the 
`rasters_reduced$dem` to:

* <span class = "subscore">[0.25]</span> Calculate the slope of each pixel in 
radians and assign to your global environment with the name "slope":

```{r 5_a, results = FALSE}

# Calculate the slope of each pixels

slope <-
  terra::terrain(
  rasters_reduced$dem,
  v = 'slope',
  unit = 'radians')

```

* <span class = "subscore">[0.25]</span> Calculate the aspect of each pixel in
radians and assign to your global environment with the name "aspect":

```{r 5_b, results = FALSE}

# Calculate the aspects of each pixel in radians

aspect <-
  terra::terrain(
  rasters_reduced$dem,
  v = 'aspect')

```

* <span class = "subscore">[0.25]</span> Generate a hillshade raster and assign
to your global environment with the name "hillshade":

```{r 5_c, results = FALSE}

# Generate a hillshade raster

hillshade <-
  hillshade <-
  terra::shade(
    slope,
    aspect,
    angle = 45,
    direction = 135)

```

* <span class = "subscore">[0.25]</span> Make a map of the topography with a 
hillshade to show vertical relief:

```{r 5_d, results = FALSE}

# Make a map of the topography with a hillshade

hillshade %>% 
  tm_shape() +
  tm_raster(
    palette = 
      gray.colors(
        n = 7, 
        
        # set the gray color from black to white
        
        start = 0, 
        end = 1), 
    style = 'cont',
    
    # mess around with the transparency as 0.9
    
    alpha = 0.9, 
    legend.show = FALSE) +
  
  # add the rasters_reduced$dem layer to show the elevation
  
  rasters_reduced$dem %>% 
  tm_shape(name = 'Elevation') +
  tm_raster(
    title = 'Elevation (m)',
    palette = terrain.colors(n = 10),
    alpha = .7) 

```

6\. <span class = "score">[1.0]</span> Nobody wants Alan to be chilly. Nobody.
We'd better help him figure out where he'll need to wear his monogrammed
Comfy&reg; fleece poncho during his trip:

* <span class = "subscore">[0.75]</span> Using average December temperatures of 
2.5 &deg;C as a cutoff, classify Alan's campsites (`camp_points`) as "Comfy" 
(temperatures less than 2.5 &deg;C) or "Flannel" (temperatures greater than or 
equal to 2.5 &deg;C);
* <span class = "subscore">[0.25]</span> Generate a tmap of Alan's potential
`campsites` where the points are colored blue for "Comfy" sites and red for
"Flannel" sites.


```{r 6, results = FALSE}

# Classify campsites as 'Comfy' or 'Flannel' using the temperature in Dec

camp_points <-
  campsites %>% 

# camp_points <- 
#   campsites %>% 
#   
#   # convert the sf object to a Spatvector
#   
#   terra::vect() %>% 
#   
#   # extract the temperature information from the Spatraster
#   
#   terra::extract(
#     rasters_reduced$tmin,
#     .) %>% 
#   
#   # classify the campsites by the temperature extracted
#   
#   mutate(
#     wear = if_else(
#       camp_points[2] < 2.5,
#       'Comfy',
#       'Flannel'))

# Generate a tmap of these potential campsites with appropriate wearings

# add a base map

tm_basemap(
  c('Esri.WorldTopoMap',
    'OpenStreetMap',
    'Esri.WorldImagery')) +

# add potential campsites with appropriate wearings
  
  # join the generated classified campsites with the geometries
  
campsites %>% 
  cbind(ID = 1:15) %>% 
  inner_join(
    camp_points) %>% 
  
  # generate a tmap and classify different dots with different wearings 
  # with blue and red
  
  tm_shape() +
  tm_dots(col = 'wear',
          palette = 
            c(Comfy = 'blue', Flannel = 'red'))
  
```

7\. <span class = "score">[1.0]</span> On long car rides, Alan likes to 
entertain his friends by singing "Don't Go Chasing Waterfalls sha-na-na-na" 
over-and-over again (punctuated only by shouting "second verse, same
as the first!"). *Waterfalls* by TLC might be Alan’s favorite song, but these are not
words that he lives by — he loves chasing waterfalls. Transylvania County, which
is one of four counties that share the Pisgah National Forest, is known as the
"Land of the Waterfalls" (it has 250 large waterfalls and many more small ones).
Let's help him find some:

* <span class = "subscore">[0.2]</span> Calculate the slope of
`rasters_reduced$dem` in degrees;
* <span class = "subscore">[0.2]</span> Mask the slope to raster cells that 
overlay `wnc_streams`;
* <span class = "subscore">[0.2]</span> Convert the raster cells to an sf point 
point file;
* <span class = "subscore">[0.2]</span> Using slopes that are greater than or
equal to 20 degrees as your criteria for a potential waterfall, subset the
resultant object to points that fit this criteria;
* <span class = "subscore">[0.2]</span> Generate a nice (but busy!) map of
waterfall points for Alan.

```{r 7_b-c, message = FALSE, warning = FALSE}

# Find waterfalls:

  # calculate the slope of rasters_reduced$dem in degrees

rasters_reduced$dem %>% 
  terra::terrain(
    v = 'slope',
    unit = 'degrees') %>% 
  
  # Mask the slope to raster cells that overlay wnc_streams
  
  terra::mask(wnc_streams %>% 
                terra::vect() %>% 
                terra::rast(resolution = c(30, 30)) %>% 
                terra::resample(rasters_reduced$dem)) %>% 
  
  # convert the raster cells to an sf point file
  
  terra::as.points() %>% 
  st_as_sf()

```

8\. <span class = "score">[1.5]</span> Sometimes ol' Alan likes to sleep out
under the stars to search for UFOs and contemplate the universe. Unfortunately,
the air in the mountains can be pretty polluted due to a coal-fired power plant
in the valley and atmospheric inversions (phenomena that trap air at low
altitudes). Let’s figure out which campsite will provide Alan with the best
chance to spot a UFO using the following criteria:

* <span class = "subscore">[0.75]</span> The campsite is 5,000 feet or more above
sea level;
* <span class = "subscore">[0.75]</span> The campsite provides a clear view of 
the night sky (canopy cover <= 10%).

*Note:*

* *You may use up to two assignments in answering this question ... but **0.25 
points extra credit** if you don't!*
* *Please return the answer as a single value that represents the name or names
of the campsite(s) (i.e., not a data frame).*


```{r 8_a-b, message = FALSE, warning = FALSE}

# Find a campsite with a clear view:

campsites %>% 
  st_coordinates()

```

9\. <span class = "score">[1.0]</span> Alan's never been scared of anything -- 
except for black bear (*Ursus americanus*). "There's just something goofy about
them ... I  don't trust goofy" Alan grunted while swirling his scotch in a way
that's *classic Alan*. Help your fellow student steer clear of black bear by
generating a density map of black bear in the area.

*Hint: Use the `bears` and `pisgah_region` shapefiles in answering this question.*

```{r 9, message = FALSE, warning = FALSE}

# Generate a density map of black bear in the area

  # create an 'owin' object

region_owin <-
  spatstat.geom::as.owin(pisgah_region)

  # generate a map

bears_ppp <-
  spatstat.geom::ppp(
    x = {bears %>% st_coordinates()}[, 1],
    y = {bears %>% st_coordinates()}[, 2], 
    window = region_owin) %>% 
  
  # convert to a density
  
  stats::density(bw = 'nrd') %>% 
  
  # generating a raster
  
  stars::st_as_stars() %>% 
  
# Plot the data:
  
  tm_shape() %>% 
  tm_raster(style = 'kmeans',
            n = 10,
            palette = 'YlGn')

```

<br>
10\. <span class = "score">[2.0]</span>. Alan's spent weeks listening intently
to Brian prattling on in this class. "How could anyone make so many verbal 
typos?" he asks himself with a long draw from his trademark Gandolf pipe. With
the course coming to a close, he's looking forward to some real peace and quiet.
Help Alan get some respite from the din by generating a map that displays the 
distance from the nearest major road (as a continuous surface).

*Hint: Depending on your choice of target raster, this may take a **long time** 
to run. As such, I strongly suggest using `rasters_reduced$precip` or 
`rasters_reduced$tmin` as your target raster!*


```{r question_10, message = FALSE, warning = FALSE}

# Paste your code in here!

```
</div>

**Extra credit** <span class = "score">[0.25]</span>:  Using data from this 
assignment, make a super impressive static map of the variable or variables of 
your choosing!

```{r extra_credit, message = FALSE, warning = FALSE}

# Paste your code in here!

```

<hr>

<script>
var acc = document.getElementsByClassName("accordion");
var i;

for (i = 0; i < acc.length; i++) {
  acc[i].addEventListener("click", function() {
    this.classList.toggle("active");
    var panel = this.nextElementSibling;
    if (panel.style.display === "block") {
      panel.style.display = "none";
    } else {
      panel.style.display = "block";
    }
  });
}
</script>