---
title: "Quiz 2: Food deserts"
output: html_document
---

<!-- Please knit this right away! The additional information for (html and css)
are provided to help you build on your current R Markdown toolbox. You will not
be responsible for learning these additional tools at this point. -->

<!-- The head tag, <head>, is a container for metadata in an html document. 
We can use it to define styles and do lots of other cool things things. -->
<head>
<!-- The link tag, <link>, creates a relationship between this file and
an external source. Here, I'm linking font-awesome so I can include a 
couple of icons that I like (user-secret and user-circle) -->
<link 
  rel="stylesheet" 
  href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">
<!-- Here I'm linking to a javascript file that will give us access to font awesome -->
  <script src="https://kit.fontawesome.com/03064bdd6c.js" crossorigin="anonymous"></script>
</head>
<!-- Note in the above that some html tags, like <head> need to be closed 
with </head>. Others, like <link> don't need a closing tag. -->

```{css, echo = FALSE}
/* We can use the body tag to change the default text size of the whole document. */
  
body {
  font-size: 18px
}

/* We can increase the space between our list items */
  
ul, li {
  padding-top: 0.25em;
}

pre {
  margin-top: 2em;
  margin-bottom: 2em;
}

/* I can add custom classes with a period symbol. */

.score {
  font-weight: bold;
  color: red;
  background-color: #ffff80;
}

/* I create an custom class for user secrets. Have a look at the area surrounding my spy icons below to see what this does. Also, do you notice any similarity to your ggPlot theme elements? */
  
.mysecret {
  border: 1px solid #999999;
  border-style: solid; 
  padding: 20px; 
  border-radius: 25px;
  margin-top: 15px;
  margin-bottom: 15px;
}

.myhint {
 color: blue; 
 font-size: 0.75em;
 font-style: italic;
}

/* Here, I add a custom class for area surrounding the graded portions of this assignment.*/

.now_you {
  border-left: 8px solid #292093;
  background-color: #e6ffff;
  padding: 16px;
  margin-top: 15px;
  margin-bottom: 15px;
}

/* Here I will add some custom css on my font awesome icons*/
  
.fas {
  font-size:36px
}
  
.subscore {
  font-size: 14px;
  color: red;
}

/* I can modify a current class like the below. To see what this does,
have a look at where I am using block quotes in the document.*/
  
blockquote {
  font-size: 18px;
  border-left: 8px solid #292093;
  background-color: #e6ffff;
  padding-left: 16px;
  padding-right: 16px;
  margin-top: 15px;
  margin-bottom: 15px;
}

/* Accordion style */
.accordion {
  background-color: #e6f0ff;
  color: #000000;
  cursor: pointer;
  padding: 18px;
  width: 100%;
  border-style: solid;
  border-width: 2px;
  border-color: #ffffff;
  text-align: left;
  outline: none;
  font-size: 15px;
  transition: 0.4s;
}

.active, .accordion:hover {
  background-color: #80b3ff; 
  color: #ffffff;
  font-weight: bold;
}

.panel {
  padding: 0 18px;
  display: none;
  background-color: #eeffff;
  overflow: hidden;
}
```

<!-- Here is the setup for knitr. I had to include tidyverse so I could
include a tidyverse example with my Important warning statement below-->

```{r knitr_setup, include = FALSE}
options(knitr.table.format = "html")

knitr::opts_chunk$set(error = TRUE)

library(tidyverse)
```

<hr>

## Overview
<div class = "row" style = "padding-left: 1em">
<a title = "Food desert cartoon">
  <img  src = "https://foodbankonline.org/wp-content/uploads/2020/06/Food-Desert.jpg"
        alt="Cartoon showing a fast food restaurant in a food desert"
        style = "float: right; 
                 padding-left: 1em; 
                 padding-bottom: 0.5em; 
                 padding-right: 1em; 
                 width: 50%; 
                 height: auto;">
  </img>
</a>

Food deserts describe geographic regions where residents have limited access to affordable healthy food options. In this exercise, you will calculate the distances between buildings and grocery stores across census tracts in the District of Columbia. Your job will be to generate a single map that represents the average distance between buildings and grocery stores in each census tract.

This exercise will give you an opportunity to test your *sf* and *tmap* skills! All content for this quiz is derived from the "Advanced sf" video lesson.
</div>

<div class = "row">
<hr>

## About the data

<br>
**dc_grocery.geojson**: This is a geojson multipolygon shapefile of grocery stores in DC.  Each polygon represents the spatial footprint of a given grocery store.

* Data source: <a href = "https://opendata.dc.gov/" target = "_blank">Open Data DC</a>;
* Primary key: There is no true primary key for these data;
* CRS: The datum of this file is WGS 84 and it is in a projected UTM coordinate system (zone 18N). 

**dc_census.geojson**: This is the census polygon shapefile that we've used across multiple exercises. 

* Data source: US Census Bureau (`tigris` and `tidycensus` packages);
* Primary key: `geoid`;
* CRS: The datum of this file is WGS84 and it is in a projected UTM coordinate system (zone 18N). 

**dc_buildings**: This is a point shapefile of buildings in DC. Each point represents the geographic centroid of a given building. To save on memory and processing time, the original data were subset to a random sample of 25% of the buildings in each census tract. Census tracts where the population is listed as "0" were removed.

* Data source: <a href = "https://opendata.dc.gov/" target = "_blank">Open Data DC</a>;
* Primary key: `object_id`;
* CRS: Geodedic WGS84 (i.e., EPSG 4326). 

</div>

<hr>


<div>
<button class="accordion">Functions that you may use in this exercise</button>
<div class="panel">

Outside of the code that I provided, you may use **only** the following R functions (*Note: If you are unclear on what a given function does, use `?` to view the help file!*):

* base::=
* base::(
* base::)
* base::as.numeric
* base::mean
* dplyr::full_join
* dplyr::group_by
* dplyr::mutate
* dplyr::summarize
* magrittr::%>%
* sf::st_crs
* sf::st_distance
* sf::st_join
* sf::st_transform
* sf::st_union
* tibble::as_tibble
* tmap::+
* tmap::tm_shape
* tmap::tm_polygons
* units::set_units

<span class = "myhint">**Note**: The packages dplyr, ggplot2, magrittr, readr, rlang, and tibble are all part of the tidyverse and are loaded with `library(tidyverse)`.</span>

</div>
</div>

<hr>

## The setup

Copy the code below to a script file and run it to set up your session (*Note: This section is exempt from the "Functions that you may use ..." above*):

```{r setup, eval = FALSE}
# setup -----------------------------------------------

library(sf)
library(tmap)
library(tidyverse)

# Set tmap mode

# Read in polygon shapefile of grocery stores:

dc_grocery <- 
  st_read('data/raw/shapefiles/dc_grocery.geojson')

# Read in polygon shapefile of census tracts in DC:

dc_census <-
  st_read('data/raw/shapefiles/dc_census.geojson') %>% 
  set_names(
    tolower(
      names(.)))

# Read in point shapefile of buildings in DC:

dc_buildings <- 
  st_read('data/raw/shapefiles/dc_buildings.geojson')
```

<div class = "now_you">
<i class="fas fa-user-circle" style = "font-size: 1.25em;"> </i>&nbsp; Without modifying the above or assigning an object to your global environment, generate a tmap that displays the average (mean) distance, in kilometers,
between buildings in a given census tract and their nearest grocery store.
</div>
<hr>

## Grading rubric

You may earn up to 5 points for this skill assessment. 

**Points earned include:**

* <span class = "subscore">[2.0]</span> You successfully calculated the distance between buildings and grocery stores;
* <span class = "subscore">[1.5]</span> You successfully summarized the average distance between buildings and grocery stores per census tract;
* <span class = "subscore">[1.5]</span> You were able generate a map of the census polygons using `tmap`.

**Points may be deducted from the question's total if the following criteria are not met:** 

* Only functions listed in "Functions that you may use in this exercise" are used <span class = "subscore">[-2.5]</span>
* The question is answered in a single piped statement  <span class = "subscore">[-2.5]</span>
* All code is properly indented  <span class = "subscore">[0.05/violation]</span>
* Your code follows the coding conventions described in *Best practices in modern R coding* <span class = "subscore">[0.05/ violation]</span>
* Your code is parsimonious <span class = "subscore">[0.05/violation]</span>

*Note: The maximum deduction is the total points value available*

<script>
var acc = document.getElementsByClassName("accordion");
var i;

for (i = 0; i < acc.length; i++) {
  acc[i].addEventListener("click", function() {
    this.classList.toggle("active");
    var panel = this.nextElementSibling;
    if (panel.style.display === "block") {
      panel.style.display = "none";
    } else {
      panel.style.display = "block";
    }
  });
}
</script>
